### 2018-05-26 Paper/app.tex
comments on application problem ...

The proposed framework in section~\ref{sec:method} is implemented to 
the H$_2$/O$_2$ reaction mechanism provided in~\cite{Yetter:1991}.
The mechanism comprises of 19 reactions including chain reactions,
dissociation/recombination reactions, and formation and consumption
of intermediate species: HO$_2$ and H$_2$O$_2$, as provided below
in Table~\ref{tab:kinetics}.

\begin{table}[htbp]
\renewcommand*{\arraystretch}{1.2}
\begin{center}
\begin{tabular}{ll}
\toprule
$\mathcal{R}_1$ & H + O$_2$ $\Longleftrightarrow$ O + OH \\
$\mathcal{R}_2$ & O + H$_2$ $\Longleftrightarrow$ H + OH \\
$\mathcal{R}_3$ & H$_2$ + OH $\Longleftrightarrow$ H$_2$O + H \\
$\mathcal{R}_4$ & OH + OH $\Longleftrightarrow$ O + H$_2$O \\
$\mathcal{R}_5$ & H$_2$ + M $\Longleftrightarrow$ H + H + M \\
$\mathcal{R}_6$ & O + O + M $\Longleftrightarrow$ O$_2$ + M \\
$\mathcal{R}_7$ & O + H + M $\Longleftrightarrow$ OH + M \\
$\mathcal{R}_8$ & H + OH +M $\Longleftrightarrow$ H$_2$O + M \\
$\mathcal{R}_9$ & H + O$_2$ + M $\Longleftrightarrow$ HO$_2$ + M \\
$\mathcal{R}_{10}$ & HO$_2$ + H $\Longleftrightarrow$ H$_2$ + O$_2$ \\
$\mathcal{R}_{11}$ & HO$_2$ + H $\Longleftrightarrow$ OH + OH \\
$\mathcal{R}_{12}$ & HO$_2$ + O $\Longleftrightarrow$ O$_2$ + OH \\
$\mathcal{R}_{13}$ & HO$_2$ + OH $\Longleftrightarrow$ H$_2$O + O$_2$ \\
$\mathcal{R}_{14}$ & HO$_2$ + HO$_2$ $\Longleftrightarrow$ H$_2$O$_2$ + O$_2$ \\
$\mathcal{R}_{15}$ & H$_2$O$_2$ + M $\Longleftrightarrow$ OH + OH + M \\
$\mathcal{R}_{16}$ & H$_2$O$_2$ + H $\Longleftrightarrow$ H$_2$O + OH \\
$\mathcal{R}_{17}$ & H$_2$O$_2$ + H $\Longleftrightarrow$ HO$_2$ + H$_2$ \\
$\mathcal{R}_{18}$ & H$_2$O$_2$ + O $\Longleftrightarrow$ OH + HO$_2$ \\
$\mathcal{R}_{19}$ & H$_2$O$_2$ + OH $\Longleftrightarrow$ HO$_2$ + H$_2$O \\
\bottomrule
\end{tabular}
\end{center}

\caption{Reaction mechanism for H$_2$/O$_2$~\cite{Yetter:1991}.}
\label{tab:kinetics}
\end{table}

The reaction rate for the $i^{th}$ reaction as a function of temperature
is given as follows:

\be
k_i(T) = A_iT^{n_i}\exp(-E_{a,i}/RT) 
\label{eq:rate}
\ee
%
where $A_i$ is the pre-exponent, $n_i$ is the index of $T$, $E_{a,i}$
is the activation energy corresponding to the $i^{th}$ reaction, and
$R$ is the universal gas constant. We focus our attention on 
quantifying the uncertainty in the `ignition delay' due to uncertainty
associated with the pre-exponent, $A_i$ for each reaction. Hence, the
total number of uncertain parameters in the present case is 19. 
While the dimensionality of the problem is relatively moderate,
constructing a surrogate in the 19D parameter space could still be
expensive. Hence, we explore the possibility of constructing a
reduce-order surrogate (ROS) using the framework presented earlier
in section~\ref{sec:method}. To this end, the $A_i$'s are considered
to be uniformly distributed in the interval: $[0.9A_i^\ast, 1.1A_i^\ast]$;
$A_i^\ast$ being the nominal estimate corresponding to the $i^{th}$
reaction. TChem~\cite{Safta:2011}, an open source C package was used
for simulating temperature and concentration evolution during the
progress of reactions ($\mathcal{R}_1$ -- $\mathcal{R}_{19}$), and determining
the ignition delay. The set of nominal values used in the computations,
for parameters in~\eqref{eq:rate} are provided in~\cite{Yetter:1991}.

In this study, we focus on two scenarios: fuel(H$_2$) rich, and fuel(H$_2$)
lean. Consider the global reaction:
%
\be
2\text{H}_2 + \text{O}_2 \rightarrow 2\text{H}_2\text{O}
\label{eq:global}
\ee 
%
The equivalence ratio ($\phi$) is defined as follows:
%
\be
\phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}}
\label{eq:phi}
\ee
%
The numerator in the RHS represents the fuel-oxygen mass ratio at a given condition
and the denominator represents the stoichiometric ratio of the same quantity. Hence,
$\phi$ = 1 at stoichiometric conditions. The equivalence ratio can be altered
by changing the amount of O$_2$ in the mixture. In the case of a lean
mixture,~\eqref{eq:global} can be written as follows:
%
\be
2\text{H}_2 + \alpha\text{O}_2 \rightarrow 2\text{H}_2\text{O} + (\alpha-1)\text{O}_2 
\hspace{3mm} (\alpha>1)
\label{eq:lean}
\ee 
%
Similarly, for the case when the mixture if fuel rich,~\eqref{eq:global} is modified
as follows:
%
\be
2\text{H}_2 + \alpha\text{O}_2 \rightarrow 2\alpha\text{H}_2\text{O} + 2(1-\alpha)\text{H}_2
\hspace{3mm} (\alpha<1)
\label{eq:rich}
\ee 
%
Eqs.~\eqref{eq:lean} and~\eqref{eq:rich} can be generalized as follows:
%
\be
2\text{H}_2 + \alpha\text{O}_2 \rightarrow 2\min(1,\alpha)\text{H}_2\text{O} + 
\max(\alpha-1,0)\text{O}_2 + \max(0,2-2\alpha)\text{H}_2
\label{eq:gen}
\ee 
%
From the above set of chemical equations, the relationship between $\phi$
and $\alpha$ can be easily obtained as $\left(\phi~=~\frac{1}{\alpha}\right)$.
Since $\phi>1$ corresponds to a rich mixture, and $\phi<1$ corresponds to a
lean mixture, we consider $\phi$ = 2.0 and 0.5 to investigate the two scenarios
respectively. The set of inputs for parameter screening: $\tau_\text{screen}$,
$s_\text{max}$, $\beta$ are fixed at 0.2, 10, and 1.0 respectively for both cases.
Additionally, the value of $\tau$ is considered to be 1.0$\times 10^{-17}$ and
5.0$\times 10^{-17}$ in the rich and lean case respectively.

A full-space surrogate based on the initial set of model evaluations at 5 samples
did not result in an accurate representation of the input-output relationship. 
Hence, unlike the test problems presented in section~\ref{sec:examples},we do not
have converged estimates of the Sobol total-effect index in this case. However,
the screening procedure in Algorithm~\ref{alg:screen} is implemented to evaluate
the screening metric for the uncertain pre-exponents ($A_i$'s) as plotted below
in Figure~\ref{fig:sense_kinetics}~(top row) for both cases. Furthermore, we illustrate
the decay in the value of $\Delta\mu_s$ with iterations in 
Figure~\ref{fig:sense_kinetics}~(bottom row).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_kinetics_rich}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_kinetics_lean}
  \\ \vspace{2mm}
  \includegraphics[width=0.48\textwidth]{./Figures/mu_rich}
  \includegraphics[width=0.48\textwidth]{./Figures/mu_lean}
\caption{Top: Estimates of $\widehat{\mathcal{C}_i\mu_i}$ for $A_i$'s in the case
of fuel-rich mixture~(left) and fuel-lean mixture~(right). Bottom: The value of
$\Delta\mu_s$ during three iterations within the screening step are plotted for
the case of fuel-rich mixture~(left) and fuel-lean mixture~(right).}
\label{fig:sense_kinetics}
\end{center}
\end{figure}
%
The screening metric estimates in the above plots are observed to converge with
only a few samples (5 -- 10). Moreover, out of the 19 uncertain pre-exponents,
only $A_1$, $A_9$, $A_{15}$, and $A_{17}$ seem to be important in the fuel-rich
case, whereas, only $A_1$, $A_9$, and $A_{15}$ seem important in the fuel-lean
case, based on the value of $\tau_\text{screen}$. These observations are indicative
of potential for severe reduction in the dimensionality of this problem. A reduced-order
surrogate, constructed using the proposed framework could thus lead to large computational
gains. The decay of $\Delta\mu_s$ with iterations is expected and provides a sanity check
for the screening procedure in both cases.

A reduced-order surrogate (ROS) was constructed in 4D for the fuel-rich case, and in 3D for
the fuel-lean case. Based on $\epsilon_{\mbox{\tiny{L-2}}}$ estimates using the validation
test suite, the ROS was found to be accurate within 1.8$\%$ in the fuel-rich case, and
within 3.1$\%$ in the fuel-lean case. Model evaluations at 1000 samples in the test suite
are further used to plot a normalized histogram of the ignition time in 
Figure~\ref{fig:pdf_kinetics}. In order to verify the accuracy of the ROS in a 
probabilistic-sense, we compare the histogram plot with a PDF of ignition time
using surrogate predictions at 10$^{6}$ samples in the reduced space in both cases. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_rich}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_lean}
\caption{A normalized histogram based on model evaluations at 1000 samples is plotted
along with a PDF of ignition delay for the fuel-rich case (left) and the fuel-lean
case (right).}
\label{fig:pdf_kinetics}
\end{center}
\end{figure}
%
Clearly, the ROS captures the spread as well as the model estimate of
the ignition delay in both scenarios. Hence, the proposed framework 
has enabled severe dimension-reduction and construction of an accurate
ROS for multiple scenarios pertaining to the H$_2$/O$_2$ reaction
mechanism.

>>>>

The proposed framework in section~\ref{sec:method} is implemented to 
the H$_2$/O$_2$ reaction mechanism provided in~\cite{Yetter:1991}.
The mechanism comprises of 19 reactions including chain reactions,
dissociation/recombination reactions, and formation and consumption
of intermediate species: HO$_2$ and H$_2$O$_2$, as provided below
in Table~\ref{tab:kinetics}.
\alennote{Are all these reactions, reversible? The notation $\Longleftrightarrow$ suggests
that. Also, 
for reversible reactions, I prefer the notation $\rightleftharpoons$.}

\begin{table}[htbp]
\renewcommand*{\arraystretch}{1.2}
\begin{center}
\begin{tabular}{llll}
\toprule
Reaction \#     & Reaction &&\\
\bottomrule
$\mathcal{R}_1$ & H + O$_2$          & $\Longleftrightarrow$ & O + OH \\
$\mathcal{R}_2$ & O + H$_2$          & $\Longleftrightarrow$ & H + OH \\
$\mathcal{R}_3$ & H$_2$ + OH         & $\Longleftrightarrow$ & H$_2$O + H \\
$\mathcal{R}_4$ & OH + OH            & $\Longleftrightarrow$ & O + H$_2$O \\
$\mathcal{R}_5$ & H$_2$ + M          & $\Longleftrightarrow$ & H + H + M \\
$\mathcal{R}_6$ & O + O + M          & $\Longleftrightarrow$ & O$_2$ + M \\
$\mathcal{R}_7$ & O + H + M          & $\Longleftrightarrow$ & OH + M \\
$\mathcal{R}_8$ & H + OH +M          & $\Longleftrightarrow$ & H$_2$O + M \\
$\mathcal{R}_9$ & H + O$_2$ + M      & $\Longleftrightarrow$ & HO$_2$ + M \\
$\mathcal{R}_{10}$ & HO$_2$ + H      & $\Longleftrightarrow$ & H$_2$ + O$_2$ \\
$\mathcal{R}_{11}$ & HO$_2$ + H      & $\Longleftrightarrow$ & OH + OH \\
$\mathcal{R}_{12}$ & HO$_2$ + O      & $\Longleftrightarrow$ & O$_2$ + OH \\
$\mathcal{R}_{13}$ & HO$_2$ + OH     & $\Longleftrightarrow$ & H$_2$O + O$_2$ \\
$\mathcal{R}_{14}$ & HO$_2$ + HO$_2$ & $\Longleftrightarrow$ & H$_2$O$_2$ + O$_2$ \\
$\mathcal{R}_{15}$ & H$_2$O$_2$ + M  & $\Longleftrightarrow$ & OH + OH + M \\
$\mathcal{R}_{16}$ & H$_2$O$_2$ + H  & $\Longleftrightarrow$ & H$_2$O + OH \\
$\mathcal{R}_{17}$ & H$_2$O$_2$ + H  & $\Longleftrightarrow$ & HO$_2$ + H$_2$ \\
$\mathcal{R}_{18}$ & H$_2$O$_2$ + O  & $\Longleftrightarrow$ & OH + HO$_2$ \\
$\mathcal{R}_{19}$ & H$_2$O$_2$ + OH & $\Longleftrightarrow$ & HO$_2$ + H$_2$O \\
\bottomrule
\end{tabular}
\end{center}
\caption{Reaction mechanism for H$_2$/O$_2$ from~\cite{Yetter:1991}}.
\label{tab:kinetics}
\end{table}

The reaction rate for the $i^{th}$ reaction as a function of temperature
is given as follows:
\be
k_i(T) = A_iT^{n_i}\exp(-E_{a,i}/RT), 
\label{eq:rate}
\ee
%
where $A_i$ is the pre-exponent, $n_i$ is the index of $T$, $E_{a,i}$ is the
activation energy corresponding to the $i^{th}$ reaction, and $R$ is the
universal gas constant. We focus on quantifying the uncertainty in the
\emph{ignition delay} due to uncertainty associated with the pre-exponent,
$A_i$ for each reaction. Hence, the total number of uncertain parameters in the
present case is 19.  The $A_i$'s are considered to be uniformly distributed in
the interval: $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ being the nominal
estimate corresponding to the $i^{th}$ reaction. TChem~\cite{Safta:2011}, an
open source C package was used for simulating temperature and concentration
evolution during the progress of reactions $\mathcal{R}_1$--$\mathcal{R}_{19}$,
and determining the ignition delay. The set of nominal values used in the
computations, for parameters in~\eqref{eq:rate} are provided
in~\cite{Yetter:1991}.

While the dimensionality of the problem is relatively moderate,
constructing a surrogate in the 19-dimensional parameter space could still be
expensive. Hence, we explore the possibility of constructing a
reduced-order surrogate (ROS) using the framework presented 
in section~\ref{sec:method}. 
In the present study, we focus on two scenarios: fuel(H$_2$) rich, and fuel(H$_2$)
lean. Consider the global reaction:
%
\be
2\text{H}_2 + \text{O}_2 \rightarrow 2\text{H}_2\text{O}
\label{eq:global}
\ee 
%
The equivalence ratio $\phi$ is defined as follows:
%
\be
\phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}}
\label{eq:phi}
\ee
%
The numerator in the right hand side represents the fuel-oxygen mass ratio at a given condition
(\alennote{What does the subscript \emph{obs} stand for?}) 
and the denominator represents the stoichiometric ratio of the same quantity. Hence,
$\phi$ = 1 at stoichiometric conditions. The equivalence ratio can be altered
by changing the amount of O$_2$ in the mixture. In the case of a lean
mixture,~\eqref{eq:global} can be written as follows:
%
\be
2\text{H}_2 + \alpha\text{O}_2 \rightarrow 2\text{H}_2\text{O} + (\alpha-1)\text{O}_2 
\hspace{3mm} (\alpha>1)
\label{eq:lean}
\ee 
%
Similarly, for the case when the mixture if fuel rich,~\eqref{eq:global} is modified
as follows:
%
\be
2\text{H}_2 + \alpha\text{O}_2 \rightarrow 2\alpha\text{H}_2\text{O} + 2(1-\alpha)\text{H}_2
\hspace{3mm} (\alpha<1)
\label{eq:rich}
\ee 
%
Eqs.~\eqref{eq:lean} and~\eqref{eq:rich} can be generalized as follows:
%
\be
2\text{H}_2 + \alpha\text{O}_2 \rightarrow 2\min(1,\alpha)\text{H}_2\text{O} + 
\max(\alpha-1,0)\text{O}_2 + \max(0,2-2\alpha)\text{H}_2
\label{eq:gen}
\ee 
%
From the above set of chemical equations, the relationship between $\phi$
and $\alpha$ can be easily obtained as $\phi~=~\frac{1}{\alpha}$.
Since $\phi>1$ corresponds to a rich mixture, and $\phi<1$ corresponds to a
lean mixture, we consider $\phi$ = 2.0 and 0.5 to investigate the two scenarios
respectively.

We apply the parameter screening algorithm with the following
parameters: $\tau_\text{screen}$,
$s_\text{max}$, $\beta$ are fixed at 0.2, 10, and 1.0 respectively for both cases.
Additionally, the value of $\tau$ is considered to be 1.0$\times 10^{-17}$ and
5.0$\times 10^{-17}$ in the rich and lean case respectively.
~\alennote{What is the rationale behind such small numbers? These are smaller
than the unit round-off.}

A full-space surrogate based on the initial set of model evaluations at 5 samples
did not result in an accurate representation of the input-output relationship. 
\alennote{This is hardly surprising, five samples in a 19D parameter space
is going to get us so far; I suggest removing or rewording this statement, 
as a careful referee will jump on this.}
Hence, unlike the test problems presented in section~\ref{sec:examples},we do not
have converged estimates of the Sobol total-effect index in this case. 
\alennote{I don't understand; did we not sample the system to death to get 
total Sobol' indices, or we are saying that's too difficult, and 
we just use DGSMs and our screening approach?}
However,
the screening procedure in Algorithm~\ref{alg:screen} is implemented to evaluate
the screening metric for the uncertain pre-exponents $A_i$'s as plotted below
in Figure~\ref{fig:sense_kinetics}~(top row) for both cases. Furthermore, we illustrate
the decay in the value of $\Delta\mu_s$ with iterations in 
Figure~\ref{fig:sense_kinetics}~(bottom row).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_kinetics_rich}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_kinetics_lean}
  \\ \vspace{2mm}
  \includegraphics[width=0.48\textwidth]{./Figures/mu_rich}
  \includegraphics[width=0.48\textwidth]{./Figures/mu_lean}
\caption{Top: Estimates of $\widehat{\mathcal{C}_i\mu_i}$ for $A_i$'s in the case
of fuel-rich mixture~(left) and fuel-lean mixture~(right). Bottom: The value of
$\Delta\mu_s$ during three iterations within the screening step are plotted for
the case of fuel-rich mixture~(left) and fuel-lean mixture~(right).}
\label{fig:sense_kinetics}
\end{center}
\end{figure}
%
The screening metric estimates in the above plots are observed to converge with
only a few samples (5--10). Moreover, out of the 19 uncertain pre-exponents,
only $A_1$, $A_9$, $A_{15}$, and $A_{17}$ seem to be important in the fuel-rich
case, whereas, only $A_1$, $A_9$, and $A_{15}$ seem important in the fuel-lean
case, based on the value of $\tau_\text{screen}$. These observations are indicative
of potential for significant reduction in the dimensionality of this problem. A reduced-order
surrogate constructed using the proposed framework could thus lead to large computational
gains. The decay of $\Delta\mu_s$ with iterations is expected and provides a sanity check
for the screening procedure in both cases.

A reduced-order surrogate (ROS) was constructed in 4D for the fuel-rich case, and in 3D for
the fuel-lean case. Based on $\epsilon_{\mbox{\tiny{L-2}}}$ estimates using the validation
test suite, the ROS was found to be accurate within 1.8$\%$ in the fuel-rich case, and
within 3.1$\%$ in the fuel-lean case. Model evaluations at 1000 samples in the test suite
are further used to plot a normalized histogram of the ignition time in 
Figure~\ref{fig:pdf_kinetics}. To verify the accuracy of the ROS in a 
probabilistic-sense, we compare the histogram plot with a PDF of ignition time
using surrogate predictions at 10$^{6}$ samples in the reduced space in both cases. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_rich}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_lean}
\caption{A normalized histogram based on model evaluations at 1000 samples is plotted
along with a PDF of ignition delay for the fuel-rich case (left) and the fuel-lean
case (right).}
\label{fig:pdf_kinetics}
\end{center}
\end{figure}
%
Clearly, the ROS captures the spread as well as the model estimate of
the ignition delay in both scenarios. Hence, the proposed framework 
has enabled significant dimension-reduction and construction of an accurate
ROS for multiple scenarios pertaining to the H$_2$/O$_2$ reaction
mechanism.
